---
title: "Case Study 5"
author: "Rachael Smith"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## (Heights) I Can Clean Your Data

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package

library(tidyverse)
library(haven)
library(readr)
library(foreign)
library(readxl)
library(downloader)
library(janitor)
library(pander)

```

```{r}

# Read in Data

Height <- read_excel("../Height.xlsx", skip = 2)
Height2 <- remove_empty(Height, which = c("rows", "cols"), quiet = TRUE)

```

### Part 1

```{r tidy_data}
# Use this R-Chunk to clean & wrangle your data!

long_height2 <- Height2 %>%
  rename("country" = `Continent, Region, Country`) %>%
  pivot_longer(cols = c(3:20), names_to = "decade2", values_to = "height.cm") %>%
  mutate(height.cm = as.numeric(height.cm))

long_height2$year_decade = long_height2$decade2
long_height2$height.in = long_height2$height.cm * 0.3937

Heights_Nation <- long_height2 %>%
  separate(decade2, into = c("century", "decade", "year"), sep = c(2, 3, 4)) %>% 
  select(Code, country, century, decade, year, height.cm, year_decade, height.in)

```


```{r plot_data}
# Use this R-Chunk to plot & visualize your data!
germ_height <- Heights_Nation %>%
  filter(country == "Germany")

ggplot(data = germ_height, mapping = aes(x = year_decade, y = height.in, col = country, group = country)) +
  geom_point() +
  geom_line() +
  labs(x = "Growth by Decade", y = "Average Male Height (Inches)", title = "Germany Average Male Height in Inches From 1810 -1980")
```

### Part 2

```{r}
# Dataset 1 G18

German_dta <- read_dta('https://byuistats.github.io/M335/data/heights/germanconscr.dta')

colnames(German_dta) [2] <- "birth_year"
colnames(German_dta) [3] <- "height.cm"
German_dta$height.in = German_dta$height.cm * 0.3937
German_dta$study = "g18"

g18 <- German_dta %>% 
  select(birth_year, height.in, height.cm, study)

```


```{r}
#Dataset 2 B19

Bavarian_dta <- read_dta('https://byuistats.github.io/M335/data/heights/germanprison.dta')
colnames(Bavarian_dta) [2] <- "birth_year"
Bavarian_dta$birth_year = as.numeric(Bavarian_dta$birth_year)
colnames(Bavarian_dta) [4] <- "height.cm"
Bavarian_dta$height.in = Bavarian_dta$height.cm * 0.3937
Bavarian_dta$study = "b19"

b19 <- Bavarian_dta %>% 
  select(birth_year, height.in, height.cm, study)

```

```{r}
# Dataset 3


```

tmpheight <- tempfile()

download("https://byuistats.github.io/M335/data/heights/Heights_south-east.zip", tmpheight, "downloader.zip", mode = "wb")

sesw_dat <- read.dbf(tmpheight, as.is = FALSE)
str(sesw_dat)
View(sesw_dat)

```{r}
#Dataset 4 US20

Bureau_dat <- read_csv('https://github.com/hadley/r4ds/raw/master/data/heights.csv')

colnames(Bureau_dat) [2] <- "height.in"
Bureau_dat$height.cm = Bureau_dat$height.in / 0.3937
Bureau_dat$birth_year = "1950"
Bureau_dat$birth_year = as.numeric(Bureau_dat$birth_year)
Bureau_dat$study = "us20"

us20 <- Bureau_dat %>% 
  select(birth_year, height.in, height.cm, study)

```


```{r}
Wisconsin_file <- read_sav("http://www.ssc.wisc.edu/nsfh/wave3/NSFH3%20Apr%202005%20release/main05022005.sav")

Wisconsin_Height <- Wisconsin_file %>% 
  select(CASE, TYPE, DOBY, RT216F, RT216I, RE35)

Wisconsin_Height$RT216F = as.numeric(Wisconsin_Height$RT216F)
Wisconsin_Height$RT216I = as.numeric(Wisconsin_Height$RT216I)
colnames(Wisconsin_Height) [3] <- "century"
Wisconsin_Height$height.inches = Wisconsin_Height$RT216F*12
Wisconsin_Height$height.in = Wisconsin_Height$height.inches +  Wisconsin_Height$RT216I
Wisconsin_Height$decade = "19"
Wisconsin_Height$study = "w20"

Wisconsin_Height$height.cm = Wisconsin_Height$height.in/0.3937

w20 <- Wisconsin_Height %>%
  filter(height.in >= 0 & height.cm >= 0 & height.in <= 90 & RE35 == 1) %>% 
  mutate(century = cut(century, breaks = c(-12, 10, 20, 30, 40, 50, 60, 70), labels = c(10, 20, 30, 40, 50, 60, 70))) %>% 
  unite(col = "birth_year", c(decade, century), sep = "")%>%
  select(birth_year, height.in, height.cm, study)

w20$birth_year = as.numeric(w20$birth_year)

```

```{r}

alld <- bind_rows(b19, g18, us20, w20)
pander(head(alld,10))
```

```{r}

ggplot(data = alld, mapping = aes(x = birth_year, y = height.in, color = study)) +
  geom_boxplot() +
  geom_jitter(height = 0, alpha = 0.25, size = .25)+
  labs(x= "Male Growth by Decade", y = "Male Height in Inches", title = "Comparison of Male Height Over the Past Century", subtitle = "Using multiple datasets and studies", color= "Study")

```

### Part 3

I feel like this data was very interesting to work with.  This was the first time we combined several datasets of different formats into a single dataset.  I can see how using a variety of data and sources could really add to the accuracy of a project, while also increaseing the complexity and pre-processing time substantially. I think the most time consuming part was simply trying to make the data match and making sure that I wasn't manipulating the results.

Having more data helped create a more diverse picture of the overall to look at. I think because the data is coming from different sources, it also helps to know that we have a very large sample to compare from.  Overall, it appears that men have been getting taller over the years.  The boxplots show an increase in the overall male height, while the Q1 of the data has stayed fairly consistent.